Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-09 GWLDCI.tex
some typos fixed and some comments left

In \cite{wilsonloop} the first author and Amat showed that every admissible Wilson loop diagrams can be identified with a unique positroid cell, and observed that this mapping is neither one-to-one nor onto. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams is given by counting the directions \hlfix{in}{of a?} associahedra.

The second paper \cite{generalcombinatoricsII} in this series identifies precisely which positroid cell is associated to a given Wilson loop diagram, by giving an algorithm for passing from a Wilson loop diagram to the Grassmann necklace of the associated positroid cell. In this manner the authors give a direct path from Wilson loop diagram to positroid cell that circumvents the painstaking process of examining bases sets. With this identification in hand, we also prove that any Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell, and characterize the volume forms that Wilson loop diagrams associate to each positroid cell.

@@

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, the \hlfix{first author and Amat}{At other points, we say "some of us" instead of specifying like this. Choose a convention} uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In (separate) papers with the second author and Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

However, the two approaches are not identical: for example, in the case of Wilson loop diagrams with $k=2$ and $n=6$, past work of some of us shows that the manifold is not contractible \cite{casestudy}. More generally, it has been shown that the spaces defined by the Amplituhedron and by Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In fact, it is in some ways more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams; in this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}.

@@

The main result of Section~\ref{sec equiv} is Theorem~\ref{same matroid iff equiv}, which says that two Wilson loop diagrams define the same matroid if and only if they are equivalent.  In subsection~\ref{sec: polygon partitions} we relate exact subdiagrams of Wilson loop diagrams to triangulations of polygons, and hence show that every Wilson loop diagram can be uniquely decomposed as a collection of exact subdiagrams (Corollary~\ref{uniqueproppartitioncor}). In subsection~\ref{sec: exact diagram matroidal props} we examine the matroidal properties of the exact subdiagrams, and prove that a subdiagram of a Wilson loop diagrams defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}). This uniform matroid is in fact equal to the dual matroid restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}).  Finally, subsection~\ref{sec: matroids and equivalence} is devoted to the proof of Theorem~\ref{same matroid iff equiv}, along with an easy corollary which gives a formula for the size of each equivalence class (Corollary~\ref{number of equiv diagrams}).

@@

\begin{dfn}\label{WLdfn}
A {\em Wilson loop diagram} is given by the following data: a cyclically ordered set $V$, along with a choice of first vertex (labeled $1$), and a set $\cP$ of $k$ pairs called {\em propagators}, written $\{p_r = (i_r, j_r)\}_{r=1}^k$. We write $W = (\cP,V)$ to denote the Wilson loop diagram. \end{dfn}

@@

Let $W = (\cP,[n])$ and $W' = (\cP',[n])$, and say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with $V(P) =  V(P')$.
\item The complementary subdiagrams are identical, i.e. $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, \hlfix{the matrix $C(W)|_V$ does not have full rank if and only if there are fewer than $|V|$ rows with non-zero entries.}{I don't think this is quite true?} This is equivalent to saying that $V$ is a set of dependent columns of $C(W)$ if and only if $|\Prop(V)| < |V|$. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, [n])$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face\footnote{That is, it is an \emph{outerplanar} graph}. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

@@

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon dissection $\tau(W)$ has a unique decomposition into maximal triangulated pieces and edges in the polygon of $\tau(W)$.
\end{lem}

@@

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees, $t$ and $t'$ in $f$ and $f'$ that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition.
This implies that $v$ has degree greater than $3$ in $T(W)$, and thus the corresponding face of $\tau(W)$ is not a triangle. In other words, the piece of $\tau(W)$ corresponding to $f'$ is not a triangulation. 
\end{proof}

@@

The previous definition makes sense, in fact, for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

\begin{cor} \label{uniqueproppartitioncor}
Any admissible Wilson loop diagram $W = (\cP, [n])$ can be uniquely decomposed into maximal exact subdiagrams. These maximal subdiagrams partition $\cP$.
\end{cor}

@@

\begin{lem} \label{lem decompose flat}Let $F$ be a flat in $M(W)$, and let $C \subseteq F$ be the union of all circuits contained in $F$. Then the following are true:
\begin{enumerate}
\item $C = F(\Prop (C))$, i.e. $C$ is a propagator flat.
\item $F \setminus C$ is an independent set. Furthermore, $F\setminus C$ is an independent flat if and only if $F(\emptyset) = \emptyset$, that is $W$ has no non-supporting vertices.  
\end{enumerate}
\end{lem}

@@

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank; we first show that $B \cup \{w\}$ is a dependent set in $M(W)$. By Lemma \ref{lem facts about WLD matroids}, we have
\begin{equation}\label{eq:equality in lem decompose flat} |\Prop(C)| = \rk(C) = |B| = \rk(B) .\end{equation}
Since $\rk(B) \leq |\Prop(B)|$, and $B \subseteq C$ implies that  $\Prop(B) \subseteq \Prop(C)$, it follows that $\Prop(B) = \Prop(C)$. Combining this with the fact that $w \in F(\Prop(C))$, we have $\Prop(B\cup \{w\}) \subseteq \Prop(B)$. We can now say something about $\rk(B\cup \{w\})$: by Lemma \ref{lem facts about WLD matroids}, we have 
\[
\rk (B \cup \{w\}) \leq \min\{|B\cup \{w\}|, |\Prop(B\cup \{w\})|\} \leq \min\{|B|+1,|Prop(B)|\},
\]
which becomes 
\[\rk(B\cup \{w\}) \leq \min\{|B|+1,|B|\} = |B| < |B \cup\{w\}|\]
via equation \eqref{eq:equality in lem decompose flat}. Therefore $B\cup\{w\}$ is a dependent set, which implies that $B\cup \{w\}$ contains a circuit $D$. In particular we must have $w \in D$ (since $B$ was an independent set), and since we have already shown that $C \subseteq F(\Prop(C))$, we also have the chain of inclusions 
\[D \subseteq B\cup \{w\} \subseteq F(\Prop(C)) \subseteq F.\]
Since $C$ was the union of all circuits in $F$, we finally obtain $w \in C$, which is a contradiction. This completes the proof of (1).

@@

\begin{cor} \label{classifyflats}
  If $F$ is a flat of an admissible Wilson loop diagram, it can be written as the disjoint union of a cyclic propagator flat and an independent set: \bas F =C \sqcup S \eas where $C$ the union of all circuits in $F$. \end{cor}

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $F(P^c)$ is a dependent flat.
\item If $(P,V(P))$ is a maximal exact subdiagram in $W$, then $F(P^c)$ is a cyclic flat.
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate}
\end{lem}

@@

To prove the result in the general case, let $(R,V(R))$ be an exact subdiagram of $W$ that is not maximal and let $P$ be the maximal exact subdiagram containing it. That is, $R \subsetneq P$.

@@

We prove the following claim: $W'$ is exact if and only if $V(P)$ contains no circuits $C$ with $\rk (C)< |P|$ in $(P, V(P))$. Since $\rk(M(W'))$ is bounded above by $|P|$, the result follows.

First suppose $C \subseteq V(P)$ is a circuit of rank $m < |P|$. Observe that $\Prop_{W'}(C)\subseteq P$ by definition, where the subscript to $\Prop$ specifies the diagram we are working in. By Lemma \ref{lem facts about WLD matroids} we know that ${|\Prop_{W'}(C)| = m}$; thus the set $P \setminus \Prop_{W'}(C)$ must be nonempty, and we can consider the subdiagram $W'':= (P\setminus \Prop_{W'}(C),V(P\setminus\Prop_{W'}(C))$. By the density condition on subdiagrams of admissible diagrams, we have
\[|V(P\setminus\Prop_{W'}(C))| \geq |P\setminus\Prop_{W'}(C)| + 3.\]
It is easy to verify that $V(P\setminus\Prop_{W'}(C)) \subseteq V(P)\setminus C$. Since $C \subseteq V(P)$ and $\Prop_{W'}(C) \subseteq P$, it follows from the previous inequality that
\[|V(P)| - (m+1) \geq |V(P\setminus\Prop(C))| \geq |P| - m + 3.\]
Simplifying, we obtain $|V(P)| \geq |P| + 4$, i.e. $(P,V(P))$ is not an exact diagram.

@@

\begin{dfn}
Given a Wilson loop diagram $W = (\cP, [n])$, its associated matroid $M(W)$ corresponds to positroid cells in the CW complex on $\Gr(|\cP|, n)$.
\end{dfn}

@@

Now consider the propagator flat $F(P)$, whichf is a cyclic flat of rank $|P|$ in both $M(W)$ and $M(W')$ by Lemma \ref{maxexactcomplementrank}. Let $B \subseteq F(P)$ be an independent set of rank $|P|$. Adding any element of $V(R)$ to $B$ increases its rank in both matroids since $F(P)$ is a flat and $F(P) \cap V(R) = \emptyset$, so we can lift any independent set in $F(P)$ to an independent set of both $M(W)$ and $M(W')$ by adjoining any independent set in $V(R)$. In particular, $B\sqcup U$ is a basis in both $M(W)$ and $M(W')$ for any $U \subseteq V(R)$ with $|U| = |R|$.

@@

For the converse, suppose that $M(W) = M(W') = M$. This immediately implies that $|\cP| = |\cP'|$, so to demonstrate the equivalence of $W$ and $W'$ it suffices to show that for any maximal exact subdiagram $(P,V(P))$ in $W$, there is a corresponding maximal exact subdiagram $(P',V(P'))$ in $W'$ with $V(P) = V(P')$.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.

@@

The correspondence between (weakly) admissible Wilson loop diagrams and polygon dissections shown in section \ref{sec: polygon partitions} suggests a connection to associahedra. First we need some background. There is a realization of associahedra that is charactertized by polygon partitions. In this section, we show that the number of equivalence classes of Wilson loop diagrams correspond to the number of non-parallel faces of the appropriate associahedron (Theorem \ref{thm:count }). While this section does not give an explicit formula for the number of these faces, we hope that this linkage with a combinatorially well studied object will shed light on the issue in the future.

@@

\begin{dfn}\label{def:secondary polytope}
  Let $x_1, \ldots, x_n$ define a convex $n$-gon, and let $T$ be the set of triangulations of this $n$-gon.  For each $t\in T$, define $s_t$ to be the point in $\mathbb{R}^n$ whose $i$th coordinate is the sum of the areas of all triangles of $t$ which are incident to $x_i$.  Let $A_n$ be the convex hull of the $s_t$; then $A_n$ a realization of the $n-1$ associahedron (\cite[Example 9.11]{Ziegler}).
\end{dfn}
Note that this construction works for any convex $n$-gon, but these realizations are all equally good for our purposes, and so we fix $x_1, \ldots, x_n$ defining a particular convex $n$-gon from now on.

@@

The map $\tau$ defined in Definition \ref{WLDtriangulationdfn} assigns a dissection of an $n$-gon to a Wilson loop diagram $W = (\cP, [n])$. Recall that if $|\cP| = k$, then $\tau(W)$ is a dissection with $k$ chords. Therefore, the map $\tau$ associates to each weakly admissible Wilson loop diagram an $n-3 - k$ dimensional face of the associahedron $A_n$. This is illustrated for $n=5$ on the left in Figure~\ref{associahedron5gon}.

@@

All edges of $A_n$ are polygon dissections which are only one diagonal away from being triangulations.  Such dissections consist of triangles and exactly one quadrilateral.  Consider the edges of $A_n$ bounding the faces corresponding to $\tau(W)$ and $\tau(X)$.  These edges have their one quadrilateral within the nontriangulated parts of $\tau(W)$ or $\tau(X)$, but these nontriangulated parts are the same in $\tau(W)$ and $\tau(X)$, so we get a bijection between the edges bounding the face of $\tau(W)$ and the edges bounding the face of $\tau(X)$ which matches edges with the same quadrilateral.

@@

Let $a$ be the area of the quadrilateral and let $a(i,j,k)$ be the area of the triangle with corners $x_i$, $x_j$ and $x_k$ (and similarly for other indices), and let $b_*$ denote the sum of the areas of triangles incident to vertex $*$ which are common to both triangulations. Then $s_{t_1}$ on coordinates $i$, $j$, $k$, $l$ is 
\[(a(i,j,l) + b_i, a+b_j, a(k,l,j)+b_k, a+b_l),\]
and $s_{t_2}$ on these coordinates is 
\[(a + b_i, a(j,k,i)+b_j, a+b_k, a(l,i,k)+b_l).\] 
This is illustrated in Figure~\ref{triangulate}.

>>>>

In \cite{wilsonloop} the first author and Amat showed that every admissible Wilson loop diagrams can be identified with a unique positroid cell, and observed that this mapping is neither one-to-one nor onto. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the directions in an associahedron.

The second paper \cite{generalcombinatoricsII} in this series identifies precisely which positroid cell is associated to a given Wilson loop diagram, by giving an algorithm for passing from a Wilson loop diagram to the Grassmann necklace of the associated positroid cell. In this manner we give a direct path from Wilson loop diagram to positroid cell circumventing the painstaking process of examining bases sets. With this identification in hand, we also prove that any Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell, and characterize the volume forms that Wilson loop diagrams associate to each positroid cell.

@@

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, the \hlfix{first author and Amat}{At other points, we say "some of us" instead of specifying like this. Choose a convention}\note{Response by Karen: I am the cause of the ``some of us'' and similar; it is the kind of phrasing I usually use, and I like it, but I am not strongly attached to it.} uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In (separate) papers with the second author and Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

However, the two approaches are not identical: for example, in the case of Wilson loop diagrams with $k=2$ and $n=6$, past work of some of us shows that the manifold is not contractible \cite{casestudy}. More generally, it has been shown that the spaces defined by the Amplituhedron and by Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In fact, it is in some ways more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams; in this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}. \note{Check with Cam.}

@@

The main result of Section~\ref{sec equiv} is Theorem~\ref{same matroid iff equiv}, which says that two Wilson loop diagrams define the same matroid if and only if they are equivalent.  In subsection~\ref{sec: polygon partitions} we relate exact subdiagrams of Wilson loop diagrams to triangulations of polygons, and hence show that every Wilson loop diagram can be uniquely decomposed as a collection of exact subdiagrams (Corollary~\ref{uniqueproppartitioncor}). In subsection~\ref{sec: exact diagram matroidal props} we examine the matroidal properties of the exact subdiagrams, and prove that a subdiagram of a Wilson loop diagram defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}). This uniform matroid is in fact equal to the dual matroid restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}).  Finally, subsection~\ref{sec: matroids and equivalence} is devoted to the proof of Theorem~\ref{same matroid iff equiv}, along with an easy corollary which gives a formula for the size of each equivalence class (Corollary~\ref{number of equiv diagrams}).

@@

\begin{dfn}\label{WLdfn}
A {\em Wilson loop diagram} is given by the following data: a cyclically ordered set $V$, along with a choice of first vertex (labeled $1$), and a set $\cP$ of $k$ pairs of elements of $V$ called {\em propagators}, written $\{p_r = (i_r, j_r)\}_{r=1}^k$. We write $W = (\cP,V)$ to denote the Wilson loop diagram. \end{dfn}

@@

Let $W = (\cP,[n])$ and $W' = (\cP',[n])$, and say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with $V(P) =  V(P')$.
\item The complementary subdiagrams are identical, i.e. $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.\note{I don't want to change it in case I've just misunderstood, but it seems to me that the vertex sets for the complementary subdiagrams should be $[n]$ not $V(P)^c$ and $V(P')^c$.}
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, \hlfix{the matrix $C(W)|_V$ does not have full rank if and only if there are fewer than $|V|$ rows with non-zero entries.}{I don't think this is quite true?  I agree, it could for instance have a column of 0s} This is equivalent to saying that $V$ is a set of dependent columns of $C(W)$ if and only if $|\Prop(V)| < |V|$. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

\begin{dfn}\label{WLDtriangulationdfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord \note{We were originally thinking in terms of chord diagrams hence the language ``chord'' but I think ``diagonal'' would be more natural now.  Any thoughts against before I change it?} edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, [n])$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It inherits an embedding from $W$ such that the vertices all lie on this infinite face\footnote{That is, it is an \emph{outerplanar} graph}. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

@@

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon dissection $\tau(W)$ has a unique decomposition into maximal triangulated pieces and edges in the polygon of $\tau(W)$.  Call this a \emph{maximal decomposition}.
\end{lem}

@@

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees, $t$ and $t'$ in $f$ and $f'$ respectively that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a non-leaf vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition.
This implies that $v$ has degree greater than $3$ in $T(W)$, and thus the corresponding face of $\tau(W)$ is not a triangle. In other words, the piece of $\tau(W)$ corresponding to $f'$ is not a triangulation. 
\end{proof}

@@

The previous definition makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

\begin{cor} \label{uniqueproppartitioncor}
Any admissible Wilson loop diagram $W = (\cP, [n])$ can be uniquely decomposed into maximal exact subdiagrams. These maximal subdiagrams partition $\cP$ and correspond to the maximal triangulated pieces of $\tau(W)$.
\end{cor}

@@

\begin{lem} \label{lem decompose flat}Let $F$ be a flat in $M(W)$, and let $C \subseteq F$ be the union of all circuits contained in $F$. Then the following are true:
\begin{enumerate}
\item $C = F(\Prop (C))$, i.e. $C$ is a propagator flat.
\item $F \setminus C$ is an independent set. Furthermore, $F\setminus C$ is an independent flat if and only if $F(\emptyset) = \emptyset$, that is, $W$ has no non-supporting vertices.  
\end{enumerate}
\end{lem}

@@

Now suppose there exists some $w \in  F(\Prop(C)) \setminus C$. Let $B$ be an independent subset of $C$ of maximal rank; we first show that $B \cup \{w\}$ is a dependent set in $M(W)$. By Lemma \ref{lem facts about WLD matroids}, we have
\begin{equation}\label{eq:equality in lem decompose flat} |\Prop(C)| = \rk(C) = |B| = \rk(B) .\end{equation}
Since $\rk(B) \leq |\Prop(B)|$, and $B \subseteq C$ implies that  $\Prop(B) \subseteq \Prop(C)$, it follows that $\Prop(B) = \Prop(C)$. Combining this with the fact that $w \in F(\Prop(C))$, we have $\Prop(B\cup \{w\}) \subseteq \Prop(B)$. We can now say something about $\rk(B\cup \{w\})$: by Lemma \ref{lem facts about WLD matroids}, we have 
\[
\rk (B \cup \{w\}) \leq \min\{|B\cup \{w\}|, |\Prop(B\cup \{w\})|\} \leq \min\{|B|+1,|\Prop(B)|\},
\]
which becomes 
\[\rk(B\cup \{w\}) \leq \min\{|B|+1,|B|\} = |B| < |B \cup\{w\}|\]
via equation \eqref{eq:equality in lem decompose flat}. Therefore $B\cup\{w\}$ is a dependent set, which implies that $B\cup \{w\}$ contains a circuit $D$. In particular we must have $w \in D$ (since $B$ was an independent set), and since we have already shown that $C \subseteq F(\Prop(C))$, we also have the chain of inclusions 
\[D \subseteq B\cup \{w\} \subseteq F(\Prop(C)) \subseteq F.\]
Since $C$ was the union of all circuits in $F$, we finally obtain $w \in C$, which is a contradiction. This completes the proof of (1).

@@

\begin{cor} \label{classifyflats}
  If $F$ is a flat of an admissible Wilson loop diagram, it can be written as the disjoint union of a cyclic propagator flat $C$ and an independent set $S$: \bas F =C \sqcup S \eas where $C$ the union of all circuits in $F$. \end{cor}

@@

\begin{lem} \label{maxexactcomplementrank}
Let $W = (\cP, [n])$ be an admissible Wilson loop diagram, and $P \subseteq \cP$. Then: \begin{enumerate}
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $F(P^c)$ is a dependent flat. \note{Should we emphasize that all these $P^c$ are wrt $W$ not the subdiagram?}
\item If $(P,V(P))$ is a maximal exact subdiagram in $W$, then $F(P^c)$ is a cyclic flat.
\item If $(P,V(P))$ is an exact subdiagram in $W$, then $\rk(F(P^c)) = |P^c|$.
\end{enumerate}
\end{lem}

@@

To prove the result in the general case, let $(R,V(R))$ be an exact subdiagram of $W$ that is not maximal and let $(P, V(P))$ be the maximal exact subdiagram containing it. That is, $R \subsetneq P$.

@@

We prove the following claim: $W'$ is exact if and only if $V(P)$ contains no circuits $C$ with $\rk (C)< |P|$ in $(P, V(P))$. Since $\rk(M(W'))$ is bounded above by $|P|$, this will give that  $M(W')$ is uniform.

First suppose $C \subseteq V(P)$ is a circuit of rank $m < |P|$. Observe that $\Prop_{W'}(C)\subseteq P$ by definition, where the subscript to $\Prop$ specifies the diagram we are working in. By Lemma \ref{lem facts about WLD matroids} we know that ${|\Prop_{W'}(C)| = m}$; thus the set $P \setminus \Prop_{W'}(C)$ must be nonempty, and we can consider the subdiagram $W'':= (P\setminus \Prop_{W'}(C),V(P\setminus\Prop_{W'}(C))$. By the density condition on subdiagrams of admissible diagrams, we have
\[|V(P\setminus\Prop_{W'}(C))| \geq |P\setminus\Prop_{W'}(C)| + 3.\]
It is easy to verify that $V(P\setminus\Prop_{W'}(C)) \subseteq V(P)\setminus C$. Since $C \subseteq V(P)$ and $\Prop_{W'}(C) \subseteq P$, it follows from the previous inequality that
\[|V(P)| - (m+1) \geq |V(P\setminus\Prop_{W'}(C))| \geq |P| - m + 3.\]
Simplifying, we obtain $|V(P)| \geq |P| + 4$, i.e. $(P,V(P))$ is not an exact diagram.

@@

\begin{dfn}
Given a Wilson loop diagram $W = (\cP, [n])$, its associated matroid $M(W)$ corresponds to positroid cells in the CW complex on $\Gr(|\cP|, n)$.
\end{dfn}\note{How is this a def as it doesn't define the cells?  Also, should it be cell singular}

@@

Now consider the propagator flat $F(P)$, which is a cyclic flat of rank $|P|$ in both $M(W)$ and $M(W')$ by Lemma \ref{maxexactcomplementrank}. Let $B \subseteq F(P)$ be an independent set of rank $|P|$. Adding any element of $V(R)$ to $B$ increases its rank in both matroids since $F(P)$ is a flat and $F(P) \cap V(R) = \emptyset$, so we can lift any independent set in $F(P)$ to an independent set of both $M(W)$ and $M(W')$ by adjoining any independent set in $V(R)$. In particular, $B\sqcup U$ is a basis in both $M(W)$ and $M(W')$ for any $U \subseteq V(R)$ with $|U| = |R|$.

@@

For the converse, \note{I don't want to mess with this because it was one of the delicate proofs, but why can't we just say $M(W)=M(W')$ so the set of uniform submatroids is the same for $M(W)$ and $M(W')$ so by the previous result the set of exact subdiagrams of $W$ and $W'$ is the same, and all this was inclusion preserving so the set of maximal exact subdiagrams is also the same.} suppose that $M(W) = M(W') = M$. This immediately implies that $|\cP| = |\cP'|$, so to demonstrate the equivalence of $W$ and $W'$ it suffices to show that for any maximal exact subdiagram $(P,V(P))$ in $W$, there is a corresponding maximal exact subdiagram $(P',V(P'))$ in $W'$ with $V(P) = V(P')$.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres\note{Is ``enumerates the fibres'' correct?} of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.

@@

The correspondence between (weakly) admissible Wilson loop diagrams and polygon dissections shown in section \ref{sec: polygon partitions} suggests a connection to associahedra. First we need some background. Associahedra can be defined in terms of polygon partitions. In this section, we show that the number of equivalence classes of Wilson loop diagrams correspond to the number of non-parallel faces of the appropriate associahedron (Theorem \ref{thm:count inequiv diagrams}). While this section does not give an explicit formula for the number of these faces, we hope that this linkage with a combinatorially well studied object will shed light on the issue in the future.

@@

\begin{dfn}\label{def:secondary polytope}
  Let $x_1, \ldots, x_n$ be the corners of a convex $n$-gon in $\mathbb{R}^2$, and let $T$ be the set of triangulations of this $n$-gon.  For each $t\in T$, define $s_t$ to be the point in $\mathbb{R}^n$ whose $i$th coordinate is the sum of the areas of all triangles of $t$ which are incident to $x_i$.  Let $A_n$ be the convex hull of the $s_t$; then $A_n$ a realization of the $n-1$ associahedron (\cite[Example 9.11]{Ziegler}).
\end{dfn}
Note that this construction works for any convex $n$-gon, but these realizations are all equally good for our purposes, and so we fix $x_1, \ldots, x_n$ defining a particular convex $n$-gon from now on.

@@

The map $\tau$ defined in Definition \ref{WLDtriangulationdfn} assigns a dissection of an $n$-gon to a Wilson loop diagram $W = (\cP, [n])$. Recall that if $|\cP| = k$, then $\tau(W)$ is a dissection with $k$ chords. Therefore, the map $\tau$ associates to each weakly admissible Wilson loop diagram an $n-3 - k$ dimensional face of the associahedron $A_n$. This is illustrated for $n=5$ on the left in Figure~\ref{associahedron5gon}, and with some dissections illustrated for $n=6$ on the right.

@@

All edges of $A_n$ are polygon dissections which are only one diagonal away from being triangulations.  Such dissections consist of triangles and exactly one quadrilateral.  Consider the edges of $A_n$ bounding the faces corresponding to $\tau(W)$ and $\tau(X)$.  These edges have their one quadrilateral within the nontriangulated parts of $\tau(W)$ or $\tau(X)$, but these nontriangulated parts are the same in $\tau(W)$ and $\tau(X)$, so we get a bijection between the edges of $A_n$ bounding the face of $\tau(W)$ and the edges of $A_n$ bounding the face of $\tau(X)$ which matches edges with the same quadrilateral.

@@

Let $a$ be the area of the quadrilateral and let $a(i,j,k)$ be the area of the triangle with corners $x_i$, $x_j$ and $x_k$ (and similarly for other indices), and let $b_i$ denote the sum of the areas of triangles incident to vertex $x_i$ which are common to both triangulations. Then $s_{t_1}$ on coordinates $i$, $j$, $k$, $l$ is 
\[(a(i,j,l) + b_i, a+b_j, a(k,l,j)+b_k, a+b_l),\]
and $s_{t_2}$ on these coordinates is 
\[(a + b_i, a(j,k,i)+b_j, a+b_k, a(l,i,k)+b_l).\] 
This is illustrated in Figure~\ref{triangulate}.
